Study of the effects of magnetic braking on the lithium abundances of the Sun and solar-type stars II

### 2022-12-08 project/overleaf/paper2/mblithium.tex
WIP paper II

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Rotation, magnetic braking \& Li abundances]{Study of the effects of magnetic braking on the lithium abundances of the Sun and solar-type stars}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[R. Caballero Navarro et al.]{
R. Caballero Navarro,$^{1}$\thanks{E-mail: rcaballeron@correo.ugr.es}
A. Garc\'ia Hern\'andez,$^{1,2}$\thanks{E-mail: agh@ugr.es}
A. Ayala$^{2},$\thanks{E-mail: aayala@ugr.es}
J.~C. Su\'arez$^{1,2}$\thanks{E-mail: jcsuarez@ugr.es}
\\
% List of institutions
% Affiliations should be in the format ‘Department, Institution, Street
% Address, City and Postal Code, Country’.
$^{1}$Dept. Theoretical Physics and Cosmology, University of Granada (UGR), 18071, Granada, Spain\\
$^{2}$Instituto de Astrof\'isica de Andaluc\'ia (CSIC), Glorieta de la Astronom\'ia S/N, 18008, Granada, Spain\\
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

@@

% Abstract of the paper
\begin{abstract}
\textit{Goals.} The study of lithium (Li) surface abundance in the Sun and young stellar globular clusters which are seemingly anomalous in present-day scenarios, as well as the influence of rotation and magnetic braking (MB) on its depletion during pre-main sequence (PMS) and main sequence (MS).
\newline\textit{Methods.} The effects of rotational mixing and of the rotational hydrostatic effects on Li abundances are studied by simulating several grids of PMS and MS rotating and non-rotating models. The previous effects are combined with the additional impact of the MB, (with magnetic field intensities ranging between 3.0 and 5.0 G). The data obtained from the two sets of simulations are confronted by comparing a number of parameters: Li abundances, angular velocity, effective temperature and size of convective zone.
\newline\textit{Results.} The surface Li abundance for the Sun like models at the end of the PMS and throughout the MS is decreased when rotational effects are included, i.e. the Li depletion rate for rotating models is higher than for non-rotating ones. However, this effect is attenuated when the braking effect produced by a magnetic field is present. This physical phenomena, generally neglected in stellar evolutionary models, impacts also the effective temperature ($\teff$) of the star and therefore its location in the HR diagram. The impact of MB in Li depletion is sensitive to the angular speed and this in turn on the magnetic field intensity: the higher the latter, the lower the Li destruction. A direct link between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. This result suggests that MB effect must be taken into consideration during PMS if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

\section{Introduction} \label{sec_intro}
Despite decades of theoretical efforts, a coherent explanation for the Li abundance discrepancies detected in star clusters cannot be found. In young and old clusters these differences have been documented for stars in both PMS and MS  evolutionary stages. Additionally, these theoretical models are not able to explain the abundances detected in the late stages of the MS \citep{Tschape2001}.\par

@@

Today, standard models are not able to replicate the observed values of Li abundance on stellar surfaces. This might happen because certain physical mechanisms that influence the destruction of Li are either being modelled improperly or simply not being taken into account, e.g. the magnetic braking. In this paper we study the interactions between rotation and magnetic fields when it comes to discuss the AM distribution. \par

@@

@@

@@

\begin{table}
	\centering
	\caption{Summary of adopted physics in MESA \citep[based on][]{Choi2016}.}
	\label{tab:phy_mesa}
	\begin{tabular}{ll} 
		\hline
		Parameter & Adopted prescriptions and values\\
		\hline
		Solar Abundance & $X_{\odot}=0.7154, Y_{\odot}=0.2703, Z_{\odot}=0.0142$\\
		Equation of State & OPAL+SCVH+MacDonald+HELM+PC\\
		Opacity & OPAL Type I for log T $\geq$ 4 \\ & Ferguson for logT $\leq$ 4\\
		Reaction Rates & JINA REACLIB\\
		Boundary Conditions & ATLAS12; $\tau$=100 tables + photosphere\\
		Difussion & Track \isotope[1]{H}, \isotope[2]{He}, \isotope[7]{Li}, \isotope[7]{Be}\\
		Rotation & Differential rotation at PMS \& MS\\
		Convection & MLT + Ledoux, $\alpha_{{\rm MLT}}$ = 1.82\\
		Overshoot & time-dependent, diffusive, \\ & $f_{{\rm ov,core}}=0.0160$,\\ 
		& $f_{{\rm ov,sh}}=0.0174$\\
		Semiconvection & $\alpha_{{\rm sc}}=0.1$\\
		Thermohaline & $\alpha_{{\rm th}}=666$\\
		Rotational Mixing & Include SH, ES, GSF, SSI \& DSI\\
		Magnetic Effects & Magnetic braking based on idealized \\ & monopole field\\
		Magnetic Field & B(G) variable between [3.0 - 5.0]\\
		Mass Loss & activated, $\Dot{M}_{{\rm max}} = 10^{-3} \: \msun \: yr^{-1}$\\
		Angular Moment Loss & activated, $\Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A}$\\
		\hline
	\end{tabular}
\end{table}

@@

\textbf{Notice the jagged appearance in Figure~\ref{fig:cz_vc_028_var_b_z1_new}. We suspected a numerical origin for the wriggles. In order to check it out, we computed models for all the physical options considered with higher temporal (time step) and spatial (number of shells) resolutions. For the sake of visual clarity, we illustrated the impact on the numerical solutions only for the 5G case, although the effects are the same in all cases. These are indicated as dotted lines in the figure, shifted-down just to become visible. Note that the larger the spatial resolution the smoother is the curve. The effect is less significant for the temporal resolution. The curve with 2x in both spatial and temporal resolution shows no difference with the case of 2x in only the number of shells. The 4x in spatial resolution points to the most appropriate number of shells to fine-track the evolution of the convective zone. However such high resolution implies to significantly increase the computation time, with no effect in the final conclusions of this work. Indeed, the size of the zigzag features is around $10^{-3}$,  which is far below the convection timescale and Li abundances measured in the stellar surface. It is thus save to neglect such effects by using the original time/space resolution.}

@@

\begin{figure}
	\includegraphics[trim = 15mm 10mm 15mm 10mm, clip,width=\columnwidth]{figures/cz_vc_028_var_g_z1_new.eps}
    \caption{\textbf{Similar to Figure \ref{fig:cz_vc_028_var_b} but now showing in detail the size of the convection zone in the models from ZAMS up to TAMS. As the intensity of the magnetic field increases, the size of the convective zone decreases. The first three pointed curves (below 0.26), have been artificially shifted downwards to illustrate the impact of the temporal (2x in time step and number of shells, red-dotted line) and spatial (2x and 4x number of shells, black and green-dotted lines, respectively) resolution on the numerical solution for the 5G case.}}
    \label{fig:cz_vc_028_var_b_z1_new}
\end{figure}

>>>>

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Rotation, magnetic braking \& Li abundances II]{Study of the effects of magnetic braking on the lithium abundances of the Sun and solar-type stars II}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[R. Caballero Navarro et al.]{
R. Caballero Navarro,$^{1}$\thanks{E-mail: rcaballeron@correo.ugr.es}
A. Garc\'ia Hern\'andez,$^{1,2}$\thanks{E-mail: agh@ugr.es}
J.~C. Su\'arez$^{1,2}$\thanks{E-mail: jcsuarez@ugr.es}
\\
% List of institutions
% Affiliations should be in the format ‘Department, Institution, Street
% Address, City and Postal Code, Country’.
$^{1}$Dept. Theoretical Physics and Cosmology, University of Granada (UGR), 18071, Granada, Spain\\
$^{2}$Instituto de Astrof\'isica de Andaluc\'ia (CSIC), Glorieta de la Astronom\'ia S/N, 18008, Granada, Spain\\
}

% These dates will be filled out by the publisher
\date{Accepted 2020 May 29. Received 2020 April 17; in original form 2019 November 30}

@@

% Abstract of the paper
\begin{abstract}
The study of lithium (Li) surface abundance in the Sun and young stellar globular clusters which are seemingly anomalous in present-day scenarios, as well as the influence of rotation and magnetic braking (MB) on its depletion during pre-main sequence (PMS) and main sequence (MS).
In this work, the effects of rotational mixing and of the rotational hydrostatic effects on Li abundances are studied by simulating several grids of PMS and MS rotating and non-rotating models. Those effects are combined with the additional impact of the MB (with magnetic field intensities ranging between 3.0 and 5.0 G). The data obtained from simulations are confronted by comparing different stellar parameters. The results show that the surface Li abundance for the Sun like models at the end of the PMS and throughout the MS decreases when rotational effects are included, i.e. the Li depletion rate for rotating models is higher than for non-rotating ones. This effect is attenuated when the MB produced by a magnetic field is present. This physical phenomena impacts also the star effective temperature ($\teff$) and its location in the HR diagram. The impact of MB in Li depletion is sensitive to the magnetic field intensity: the higher it is, the lower the Li destruction. A direct link between the magnetic fields and the convective zone (CZ) size is observed: stronger magnetic fields produce shallower CZ's. This result suggests that MB effect must be taken into consideration during PMS if we aim to reproduce Li abundances in young clusters.
\end{abstract}

@@

\section{Introduction} \label{sec_intro}
Over the last few years, decades, data from different observational campaigns have allowed studies of the rotational evolution of low and medium-mass stars (REFERENCES). Among them we find studies (REFERENCES) that focus on modelling the effects of magnetic braking and how its coupling with stellar winds impact the rotational evolution of the star. These models prefix a number of parameters in a fixed way, among them the magnetic field strength B (we will come back to this point later). This approach is not sufficient to explain the observational data, and interaction with other processes in the angular velocity evolution of low-mass stars is necessary: the distribution of angular momentum loss in the interior of the star along the pre-main sequence (PMS; REFERENCES), and main sequence (MS; REFERENCES), and the star - protostellar disc interaction during the PMS (REFERENCES).

Despite decades of theoretical efforts, a coherent explanation for the Li abundance discrepancies detected in star clusters cannot be found. In young and old clusters these differences have been documented for stars in both PMS and MS  evolutionary stages. Additionally, these theoretical models are not able to explain the abundances detected in the late stages of the MS \citep{Tschape2001}.\par

@@

Today, standard models are not able to replicate the observed values of Li abundance on stellar surfaces. This might happen because certain physical mechanisms that influence the destruction of Li are either being modelled improperly or simply not being taken into account, e.g. the magnetic braking. In this paper we study the interactions between rotation and magnetic fields when it comes to discuss the AM distribution. \par
---NEW---
In the first paper of this series \citep{Navarro2020}, we developed a model to account for the angular momentum evolution of solar-mass stars. The parametric model included a magnetic braking law, and mass-loss prescriptions that relied on \citet{Ud-Doula2008}, and core-envelope differential rotation at PMS and MS. The aim of the present study is to extend that model to with a law of interdependence between $\Omega$ and $B$. So that, during the PMS, when the star rotates faster, the MB effect is more efficient. On the other hand, during the MS, when the start slows down, the MB will become less intensive. This approach would establish a self-regulating mechanism over the angular velocity of the star that
would end up directly influencing the Li evolution.\par

Following a similar approach, we aim the model the evolution of alpha MLT with another law of interdependence between $Teff$ and surface gravity, both of them directly depending on $\Omega$.\par

---NEW---

\section{Data} \label{sec_method}
Mirar Gaia-ESO Survey: Role of magnetic activity and starspots on pre-main-sequence lithium evolution

@@

\subsection{Modelling magnetic intensity} \label{mod_mbi}

@@

\subsection{Modelling MLT $\alpha$} \label{mod_mltalpha}

@@

\begin{table}
	\centering
	\caption{Summary of adopted physics in MESA \citep[based on][]{Choi2016,Navarro2020}.}
	\label{tab:phy_mesa}
	\begin{tabular}{ll} 
		\hline
		Parameter & Adopted prescriptions and values\\
		\hline
		Solar Abundance & $X_{\odot}=0.7154, Y_{\odot}=0.2703, Z_{\odot}=0.0142$\\
		Equation of State & OPAL+SCVH+MacDonald+HELM+PC\\
		Opacity & OPAL Type I for log T $\geq$ 4 \\ & Ferguson for logT $\leq$ 4\\
		Reaction Rates & JINA REACLIB\\
		Boundary Conditions & ATLAS12; $\tau$=100 tables + photosphere\\
		Diffusion & Track \isotope[1]{H}, \isotope[2]{He}, \isotope[7]{Li}, \isotope[7]{Be}\\
		Rotation & Differential rotation at PMS \& MS\\
		Convection & MLT + Ledoux, variable $\alpha_{{\rm MLT}}$\\
		Overshoot deactivated & $f_{{\rm ov,core}}=0.0$, \& $f_{{\rm ov,sh}}=0.0$\\
		Semiconvection & $\alpha_{{\rm sc}}=0.1$\\
		Thermohaline & $\alpha_{{\rm th}}=666$\\
		Rotational Mixing & Include SH, ES, GSF, SSI \& DSI\\
            Mixing Length Theory & $\alpha_{{\rm MLT}}$ variable depending on $\teff$\\ & \& effective surface gravity\\
		Magnetic Effects & Magnetic braking based on idealized \\ & monopole field\\
		Magnetic Field & B(G) variable, depending on $\teff$ \&  $\Omega$\\
		Mass Loss & activated, $\Dot{M}_{{\rm max}} = 10^{-3} \: \msun \: yr^{-1}$\\
		Angular Moment Loss & activated, $\Dot{J} = \frac{2}{3} \Dot{M}\Omega R^{2}_{A}$\\
		\hline
	\end{tabular}
\end{table}

@@

Notice the jagged appearance in Figure~\ref{fig:cz_vc_028_var_b_z1_new}. We suspected a numerical origin for the wriggles. In order to check it out, we computed models for all the physical options considered with higher temporal (time step) and spatial (number of shells) resolutions. For the sake of visual clarity, we illustrated the impact on the numerical solutions only for the 5G case, although the effects are the same in all cases. These are indicated as dotted lines in the figure, shifted-down just to become visible. Note that the larger the spatial resolution the smoother is the curve. The effect is less significant for the temporal resolution. The curve with 2x in both spatial and temporal resolution shows no difference with the case of 2x in only the number of shells. The 4x in spatial resolution points to the most appropriate number of shells to fine-track the evolution of the convective zone. However such high resolution implies to significantly increase the computation time, with no effect in the final conclusions of this work. Indeed, the size of the zigzag features is around $10^{-3}$,  which is far below the convection timescale and Li abundances measured in the stellar surface. It is thus save to neglect such effects by using the original time/space resolution.

@@

\begin{figure}
	\includegraphics[trim = 15mm 10mm 15mm 10mm, clip,width=\columnwidth]{figures/cz_vc_028_var_g_z1_new.eps}
    \caption{Similar to Figure \ref{fig:cz_vc_028_var_b} but now showing in detail the size of the convection zone in the models from ZAMS up to TAMS. As the intensity of the magnetic field increases, the size of the convective zone decreases. The first three pointed curves (below 0.26), have been artificially shifted downwards to illustrate the impact of the temporal (2x in time step and number of shells, red-dotted line) and spatial (2x and 4x number of shells, black and green-dotted lines, respectively) resolution on the numerical solution for the 5G case.}
    \label{fig:cz_vc_028_var_b_z1_new}
\end{figure}
